---
title: "GRADUATE STUDENT STAT 840 A1"
author: "Vsevolod Ladtchenko 20895137"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 2

## (a)

We know that if $X \sim N(\theta, 1)$ then $\bar{X}_n \sim N(\theta,\frac{1}{n})$

Likelihood of $\bar{X}_{10}$:

$$
f(x) = \frac{\sqrt{10}}{\sqrt{2 \pi}} e^{- \frac{1}{2} \left (\sqrt{10} (x - \theta) \right )^2}
$$

```{r}
thetas = seq(from = -1, to = 6, length.out = 100)
lik_a = dnorm(3.5, thetas, 1/sqrt(10))
```

## (b)

We know $\sqrt{n} (\bar{x} - \theta) \sim N(0,1)$, which allows us to express the given condition as a probability statement:

$$
\begin{aligned}
0 &< \bar{x}_{10} < 5 \\
0 - \theta &< \bar{x}_{10} - \theta < 5 - \theta \\
 - \sqrt{10} \theta &< \sqrt{10} (\bar{x}_{10} - \theta) < \sqrt{10}(5 - \theta) \\
&\Phi(\sqrt{10}(5 - \theta)) - \Phi(- \sqrt{10} \theta)
\end{aligned}
$$

```{R}
lik_b = pnorm(sqrt(10)*(5 - thetas)) - pnorm(-sqrt(10)*thetas)
```


## (c)

For part A, the likelihood is highest when $\theta = \bar{x}$, which is the usual MLE. It decreases around 3.5 in accordance to the sample mean's variance, which is $1/\sqrt{10} \approx 0.31$

For part B, we only know that the mean is in an interval. Standardizing this inequality gives us a probability statement. If $\theta$ was $0$ then we have $50$% chance the mean is to the right, and if $\theta$ is 5, $50$% chance it is to the left. Since $\sigma = 0.31$, $3$ standard deviations is approximately $1$, we see that we quickly climb to $1$ in the interval $(0,5)$. This shows that if $\theta$ is here, then we are almost guaranteed to have $\bar{x}$ in $(0,5)$, except near the individual points $0$ and $5$. 

Naturally, having more precise information about $\bar{x}$ makes the likelihood that much more informative. 

```{R}
plot(thetas, lik_a, type='l', col='red')
lines(thetas, lik_b, col='green')
legend(x = 'topright', 
       legend=c('A', 'B'), 
       fill = c('red','green'))
```