---
title: "GRADUATE STUDENT STAT 840 A1"
author: "Vsevolod Ladtchenko 20895137"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1

## (a) + (b)

A likelihood:

$$
f_A(x) = \frac{\beta^{\alpha}}{\Gamma(\alpha)} x^{\alpha-1}e^{-\beta x} = \frac{\beta^{3}}{\Gamma(3)} x^{3-1}e^{-\beta x} = \frac{\beta^{3}}{2} x^{2}e^{-\beta x} = \frac{\theta^{3}}{2} x^{2}e^{-\theta x}
$$
$$
f_A(2) = 2 \theta^3 e^{-2\theta}
$$

B likelihood:

$$
f_B(x) = \frac{\lambda^x e^{-\lambda}}{x!} = \frac{2^x \theta^x e^{-2\theta}}{x!}
$$
$$
f_B(3) = \frac{8}{6} \theta^3 e^{-2 \theta}
$$
In particular,

$$
f_A(2) = \frac{6}{4} f_B(3)
$$

```{r}
n = 100
results1 = rep(0, n)
results2 = rep(0, n)
theta = seq(from = 0, to = 10, length.out = n)

for (i in 1:n)
{
  results1[i] = dgamma(2, 3, theta[i])
  results2[i] = dpois(3, theta[i]*2)
}

plot(theta, results1, type='l', col='red')
lines(theta, results2, col='green')
legend(x = 'topright', 
       legend=c('A gam', 'B pois'), 
       fill = c('red','green'))
```

## (c)

Comparing the likelihoods, we see that $L_A$ is proportional to $L_B$ as a function of $\theta$. The MLE will thus be the same, yielding $\theta^{\star} = 3/2$.

$$
\begin{aligned}
loglik &= c + 3 log \theta - 2 \theta \\
\frac{d}{d\theta} (c + 3 log \theta - 2 \theta) &= 0 \\
\frac{3}{\theta} &= 2 \\ 
\theta &= \frac{3}{2}
\end{aligned}
$$

### Discussion:
According to the Strong Likelihood Principle, the likelihood function contains all information $x$ has about $\theta$. If $x$, $y$ are two observations (possibly from different models) satisfying $L_A = c L_B$, for every $\theta$, they carry the same information about $\theta$ and must lead to identical inference. The constraints are satisfied in this situation, meaning both experimenters A and B have the same inference about $\theta$, given their chosen models and the observations $x$, $y$. Their likelihood functions are proportional. So A and B make identical inferences about $\theta$.