---
title: "STAT 840 Chapter 1 Exercises"
author: "Vsevolod Ladtchenko 20895137"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# TODO

[1] 1
[1] 2
[1] 3
[1] 4
[1] 5
[1] 6
[1] 7
[1] 8
[1] 9
[1] 10
[1] 11
[1] 12
[1] 13
[1] 14

# TRICKS TO REMEMBER

- mode = highest point of density. get it by taking derivative

- gamma/beta mean = rewrite as Gam(a+1,b) using Gam(n) = (n-1)!

- chap 1 variance decomposition: use only decomposition for xbar

# PROBABLY CORRECT BUT I SHOULD ASK ANYWAYS

18

# Problem 1.14


# Problem 1.15

$$
\begin{aligned}
f(x \mid \theta) &\propto \theta^x (1 - \theta)^{1-x} \\ 
f(X \mid \theta) &\propto \prod_{i=1}^n \theta^{x_i} (1 - \theta)^{1-x_i} \\
&\propto  \theta^{\sum_{i=1}^n x_i} (1 - \theta)^{n- \sum_{i=1}^n x_i} \\
f(\theta) &\propto \theta^{\alpha-1} (1-\theta)^{\beta-1} \\
f(\theta \mid X) &= f(X \mid \theta) f(\theta) \\
&\propto \theta^{\sum_{i=1}^n x_i} (1 - \theta)^{n- \sum_{i=1}^n x_i} \theta^{\alpha-1} (1-\theta)^{\beta-1} \\
&\propto \theta^{\alpha-1+\sum_{i=1}^n x_i} (1 - \theta)^{\beta-1+n- \sum_{i=1}^n x_i} \\
&\sim \text{Beta}(\alpha_{prior} + \sum_{i=1}^n x_i, \beta_{prior}+n- \sum_{i=1}^n x_i) \\
\end{aligned}
$$

```{r}
problem_1_15 = function()
{
  x = c(0, 1, 0, 1, 0, 0, 0, 0, 1, 0, 0, 0, 0)
  
  bay_est = function (a,b) a/(a+b)
  
  map_est = function (a,b) (a-1)/(a+b-2)
  
  params = c(100, 10, 1, 0.5)
  colors = c('red','green','blue','purple')
  
  for (i in 1:4)
  {
    a_prior = params[i]
    b_prior = params[i]
    
    a_post = a_prior + sum(x)
    b_post = b_prior + length(x) - sum(x)
    
    x_ = seq(0,1,1/1000)
    y_ = dbeta(x_, a_post, b_post)
    
    if (i==1) 
    {
      plot(x=x_,y=y_,type='l', col=colors[i])
      legend(x = "topright", legend=params, fill = colors)
    }
    else lines(x=x_,y=y_, col=colors[i])
    
    print(paste("for prior (", a_prior, ",",b_prior, ")"))
    print(round(c(Bay=bay_est(a_post, b_post), MAP=map_est(a_post, b_post)),4))
  }
}

problem_1_15()
```

# Problem 1.16

$$
\begin{aligned}
X_i &\sim Pois(\theta) \\
f(x_i \mid \theta) &= \frac{\theta^{x_i} e^{-\theta}}{x_i!} \\ 
f(X \mid \theta) &= \prod_{i=1}^n \frac{\theta^{x_i} e^{-\theta}}{x_i!} \\
f(\theta) &= \frac{\beta^\alpha}{\Gamma(\alpha)} \theta^{\alpha-1} e^{-\beta \theta} \\
f(\theta \mid X) &= f(X \mid \theta) f(\theta) \\
&=  \frac{\beta^\alpha}{\Gamma(\alpha)} \theta^{\alpha-1} e^{-\beta \theta} \prod_{i=1}^n \frac{\theta^{x_i} e^{-\theta}}{x_i!} \\
&\propto  \theta^{\alpha-1} e^{-\beta \theta} \prod_{i=1}^n \theta^{x_i} e^{-\theta} \\
&\propto  \theta^{\alpha-1} e^{-\beta \theta -n \theta} \theta^{\sum_{i=1}^n x_i} \\
&\propto  \theta^{\alpha-1 + \sum_{i=1}^n x_i} e^{-\beta \theta -n \theta} \\
&\propto  \theta^{(\alpha + \sum_{i=1}^n x_i) - 1} e^{- \theta (\beta + n) } \\
&\propto  \theta^{(\alpha + \sum_{i=1}^n x_i) - 1} e^{- (\beta + n) \theta } \\
&\sim \text{Gamma}(\alpha_{post} = \alpha_{prior} + \sum_{i=1}^n x_i, \beta_{post} = \beta_{prior} + n) \\
\end{aligned}
$$

Posterior mode:

$$
\begin{aligned}
f(\theta) &= \frac{\beta^\alpha}{\Gamma(\alpha)} \theta^{\alpha-1} e^{-\beta \theta} \\
f(\theta) &\propto \theta^{\alpha-1} e^{-\beta \theta} \\
f^\prime(\theta) &\propto \theta^{\alpha-1} (-\beta)e^{-\beta \theta} + (\alpha-1) \theta^{\alpha-2} e^{-\beta \theta} \\
0 &= \theta^{\alpha-1} (-\beta) + (\alpha-1) \theta^{\alpha-2} \\
0 &= (-\beta\theta  + \alpha-1) \theta^{\alpha-2} \\
& \qquad \text{assume } \theta \neq 0 \text{ and later check value at } 0\\
0 &= -\beta\theta  + \alpha-1 \\
\beta\theta  &= \alpha-1 \\
\theta  &= \frac{\alpha-1}{\beta}\\
& \qquad \text{we know support is positive} \\
& \qquad \text{this is positive for } \alpha > 1 \\
& \qquad \alpha < 1 \text{ implies max at 0} \\
\end{aligned}
$$

Posterior mean:

$$
\begin{aligned}
f(\theta) &= \frac{\beta^\alpha}{\Gamma(\alpha)} \theta^{\alpha-1} e^{-\beta \theta} \\
\mathbb{E} [ \theta] &= \int_0^\infty \theta \frac{\beta^\alpha}{\Gamma(\alpha)} \theta^{\alpha-1} e^{-\beta \theta} d\theta \\
&= \int_0^\infty \frac{\beta^\alpha}{\Gamma(\alpha)} \theta^{(\alpha+1)-1} e^{-\beta \theta}  d\theta \\
&= \int_0^\infty \frac{1}{\beta} \frac{\beta^{\alpha + 1 }}{\Gamma(\alpha)} \theta^{(\alpha+1)-1} e^{-\beta \theta}  d\theta \\
& \qquad \Gamma(\alpha) = (\alpha - 1)! \\
& \qquad \Gamma(\alpha+1) = \alpha! \\
& \qquad \Gamma(\alpha+1) = \alpha \Gamma(\alpha) \\
&= \int_0^\infty \frac{\alpha}{\beta} \frac{\beta^{\alpha + 1 }}{\Gamma(\alpha + 1)} \theta^{(\alpha+1)-1} e^{-\beta \theta}  d\theta \\
&= \frac{\alpha}{\beta}  \int_0^\infty \frac{\beta^{\alpha + 1 }}{\Gamma(\alpha + 1)} \theta^{(\alpha+1)-1} e^{-\beta \theta}  d\theta \\
&= \frac{\alpha}{\beta} \\
\end{aligned}
$$

Thus the posterior mean and mode of $\theta$ are:

$$
\begin{aligned}
\theta_{mode} &= \frac{\alpha_{prior} + \sum_{i=1}^n x_i-1}{\beta_{prior} + n} \\
\theta_{mean} &= \frac{\alpha_{prior} + \sum_{i=1}^n x_i}{\beta_{prior} + n}
\end{aligned}
$$

# Problem 1.17

$$
\begin{aligned}
\theta &\sim \text{Beta}(\alpha=5, \beta=10) \\
f(\theta) &\propto \theta^{\alpha-1} (1-\theta)^{\beta-1}\\
&\propto \theta^{4} (1-\theta)^{9}\\
f(x \mid \theta) &= {n \choose k} \theta^k (1 - \theta)^{n-k} \\
&\propto \theta^1 (1 - \theta)^{19} \\
f(\theta \mid x) &= f(x \mid \theta) f(\theta) \\
&\propto \theta^1 (1 - \theta)^{19} \theta^{4} (1-\theta)^{9} \\
&\propto \theta^5 (1 - \theta)^{28} \\
&\sim \text{Beta}(6, 29)
\end{aligned}
$$

MAP estimate (mode):

$$
\begin{aligned}
f(\theta \mid x) &\propto \theta^{\alpha-1} (1 - \theta)^{\beta-1} \\
f^\prime(\theta \mid x) &\propto -(\beta-1)\theta^{\alpha-1} (1 - \theta)^{\beta-2} + (\alpha-1) \theta^{\alpha-2} (1 - \theta)^{\beta-1}\\
0 &= \Biggl( -(\beta-1)\theta  + (\alpha-1) (1 - \theta) \Biggr) \theta^{\alpha-2} (1 - \theta)^{\beta-2}\\
& \qquad \text{assume } \theta, (1 - \theta) \neq 0 \text{ later check value at } 0 \\
0 &= -(\beta-1)\theta  + (\alpha-1) (1 - \theta) \\
0 &= (1 -\beta)\theta  + \alpha(1 - \theta) - (1 - \theta)  \\
0 &= \theta -\beta\theta  + \alpha - \alpha \theta - 1 + \theta  \\
1 -\alpha &= 2 \theta - \alpha \theta -\beta\theta \\
1 -\alpha &= (2 - \alpha -\beta)\theta \\
\theta &= \frac{1 -\alpha}{2 - \alpha -\beta} \frac{-1}{-1} \\
\theta &= \frac{\alpha - 1}{\alpha + \beta -2} \\
\end{aligned}
$$

Bayes estimate (mean):

$$
\begin{aligned}
f(\theta \mid x) &= \frac{\Gamma(\alpha + \beta)}{\Gamma(\alpha) \Gamma(\beta) } \theta^{\alpha-1} (1 - \theta)^{\beta-1} \\
\mathbb{E}[\theta] &= \int_0^1 \theta \frac{\Gamma(\alpha + \beta)}{\Gamma(\alpha) \Gamma(\beta) } \theta^{\alpha-1} (1 - \theta)^{\beta-1} d\theta \\
&= \int_0^1 \frac{\Gamma(\alpha + \beta)}{\Gamma(\alpha) \Gamma(\beta) } \theta^{(\alpha+1)-1} (1 - \theta)^{\beta-1} d\theta \\
& \qquad \Gamma(\alpha+1) = \alpha \Gamma(\alpha) \\
& \qquad \Gamma(\alpha+\beta + 1) = (\alpha+\beta) \Gamma(\alpha + \beta) \\
&= \int_0^1 \frac{\alpha}{\alpha+\beta} \frac{\Gamma(\alpha +1 + \beta)}{\Gamma(\alpha+1 ) \Gamma(\beta) } \theta^{(\alpha+1)-1} (1 - \theta)^{\beta-1} d\theta \\
&= \frac{\alpha}{\alpha+\beta} \int_0^1 \frac{\Gamma(\alpha +1 + \beta)}{\Gamma(\alpha+1 ) \Gamma(\beta) } \theta^{(\alpha+1)-1} (1 - \theta)^{\beta-1} d\theta \\
&= \frac{\alpha}{\alpha+\beta} \\
\end{aligned}
$$

So in our problem:

$$
\begin{aligned}
\theta_{mode} &= \frac{6 - 1}{6 + 29 -2} \\
\theta_{mean} &= \frac{6}{6+29}  \\
\end{aligned}
$$

# Problem 1.18

$$
\begin{aligned}
P(\text{positive}) &= P(\text{positive} \cap \text{sensitive}) + P(\text{positive} \cap \text{not-sensitive}) \\
& \qquad \text{law of total probability} \\
&= P(\text{positive} \mid \text{sensitive}) P(\text{sensitive}) + P(\text{positive} \mid \text{not-sensitive}) P(\text{not-sensitive}) \\
& \qquad \text{definition of conditional probability} \\
&= P(\text{positive} \mid \text{sensitive})0.5 + P(\text{positive} \mid \text{not-sensitive}) 0.5 \\
&= p0.5 + 0.25 \\
0.5 p &= P(\text{positive}) - 0.25 \\
p &= 2 P(\text{positive}) - 0.5 \\
\end{aligned}
$$

Now we need to get the MLE of the probability of a positive response. Since this involves counting a group of binary outcomes, it is a binomial distribution.

$$
\begin{aligned}
f(k) &= {n \choose k} \theta^k (1 - \theta)^{n-k} \\
\frac{d}{d \theta} f(k) &= -(n-k){n \choose k} \theta^k (1 - \theta)^{n-k-1} + k {n \choose k} \theta^{k-1} (1 - \theta)^{n-k} \\
0 &= \Biggl( -(n-k) \theta  + k  (1 - \theta) \Biggr) \theta^{k-1} (1 - \theta)^{n-k-1} \\
& \qquad \text{assume } \theta, (1 - \theta) \neq 0 \text{ later check value at } 0 \\
0 &= -(n-k) \theta  + k  (1 - \theta) \\
0 &= (k-n) \theta  + k - k\theta \\
0 &= k\theta - n\theta  + k - k\theta \\
0 &= - n\theta  + k \\
n\theta &= k \\
\theta &= k/n \\
\end{aligned}
$$

Because of the invariance property of the MLE, we know that any function of the MLE is the MLE of that function. So in our case:

$$
\begin{aligned}
P(\text{positive})_{MLE} &= X/n \\
p_{MLE} &= 2 P(\text{positive})_{MLE} - 0.5 \\
p_{MLE} &= 2 X/n - 0.5 \\
\end{aligned}
$$

# Problem 1.19

## i)

$$
\begin{aligned}
\sum_{i=1}^n (x_i - \bar{x}_n)^2 &= \sum_{i=1}^n x_i^2 - 2 \bar{x}_n \sum_{i=1}^n x_i + \sum_{i=1}^n \bar{x}_n^2 \\
&= \sum_{i=1}^n x_i^2 - 2 \bar{x}_n n \bar{x}_n  + n \bar{x}_n^2 \\
& \qquad \text{since } n \bar{x}_n = \sum_{i=1}^n x_i \\
&= \sum_{i=1}^n x_i^2 - n \bar{x}_n^2  \\
\end{aligned}
$$

## ii)

$$
\begin{aligned}
\bar{x}_{n+1} &= \frac{1}{n+1}\sum_{i=1}^{n+1} x_i \\ 
&= \frac{1}{n+1} x_{n+1} + \frac{1}{n+1}\sum_{i=1}^{n} x_i  \\ 
&= \frac{1}{n+1} x_{n+1} + \frac{n}{n+1} \bar{x}_{n} \\ 
&= \bar{x}_{n} + \frac{1}{n+1} (x_{n+1}- \bar{x}_{n})\\ 
\end{aligned}
$$

$$
\begin{aligned}
\sum_{i=1}^{n+1} (x_i - \bar{x}_{n+1})^2 &= \sum_{i=1}^{n+1} (x_i - \bar{x}_{n} - \frac{1}{n+1} (x_{n+1}- \bar{x}_{n}))^2 \\
& \qquad \text{from previous part, plug in } \bar{x}_{n+1} \\
&= \sum_{i=1}^{n+1} (x_i - \bar{x}_{n})^2   - 2 \frac{1}{n+1} (x_{n+1}- \bar{x}_{n}) \sum_{i=1}^{n+1} (x_i - \bar{x}_{n}) + \sum_{i=1}^{n+1} \frac{1}{(n+1)^2} (x_{n+1}- \bar{x}_{n})^2 \\
& \qquad \text{note: }\sum_{i=1}^{n+1} (x_i - \bar{x}_{n}) = (n+1)(\bar{x}_{n+1} - \bar{x}_{n}) \\
& \qquad \text{from previous part, } \bar{x}_{n+1} - \bar{x}_{n} = \frac{1}{n+1} (x_{n+1}- \bar{x}_{n}) \\
&= \sum_{i=1}^{n+1} (x_i - \bar{x}_{n})^2   - \frac{2}{n+1} (x_{n+1}- \bar{x}_{n})^2 + \frac{1}{n+1} (x_{n+1}- \bar{x}_{n})^2 \\
&= \sum_{i=1}^{n} (x_i - \bar{x}_{n})^2  + \frac{n}{n+1} (x_{n+1}- \bar{x}_{n})^2 \\
\end{aligned}
$$


